### 2013-12-04 rebuttal.txt
Manuscript revised based on reviewer comments
- Tried to address all reviewer concerns
- Also included responses to reviewer comments in rebuttal.txt

Dear Editor,

Please find a revised version of our Methods paper "Nengo: A Python
tool for building large-scale functional brain models" attached for
your consideration for publication in Frontiers in Neuroinformatics.

Sincerely,

Trevor Bekolay, James Bergstra, Eric Hunsberger, Travis DeWolf,
Terrence C. Stewart, Daniel Rasmussen, Xuan Choo, Aaron Voelker, and
Chris Eliasmith

Reviewer 1
==========

Expert Authority
----------------

>> Are relevant methods of other authors mentioned? If not, please
>> specify the methods the author should incorporate.

> Emergent should be mentioned. It is not based on spiking simulators,
> but has a track record in cognitive modelling.

We have now cited Emergent in the introduction alongside ACT-R in the
introduction.

> MIIND could be mentioned (Disclaimer: the reviewer is involved with
> MIIND) It models population dynamics, admittedly not based on
> spiking neurons, but again its aim is large-scale cognitive models
> and shares some design ideas (see paper in Neural Networks). Please
> judge for yourself.

We have revised Section 7.1, which now includes a list of tools
that simulate cognitive phenomena, including MIIND.

The revised text reads:

"There are many other neural simulators dedicated to building
large-scale neural models (Eppler et al., 2008; Goodman and Brette,
2009; Hines et al., 2009), and many tools for simulating cognitive
phenomena with various levels of biologically plausibility (Cooper and
Fox, 1998; Sun, 2001; Anderson et al., 2004; Franklin et al., 2007;
Aisa et al., 2008; de Kamps et al., 2008; Laird, 2012). However, Nengo
is unique in being built on a theoretical framework that has enabled a
cognitive architecture (the Semantic Pointer Architecture) that
maintains a high level of biological plausibility, and has been
validated through the Spaun model and other past work."

Other Comments
--------------

>> Please add any further comments you have regarding this manuscript.

> It looks like an interesting tool, and I think at this level
> existing simulators can do with a bit of competition. But the
> description of the tool leaves much to be desired. In particular it
> doesn't make clear enough what its advantages over NEST or BRIAN
> are. The discussion on scaling is not entirely convincing because in
> part, it is hard to understand how to compare with those simulators.

We have expanded Section 7.1 to compare to projects that have similar
design goals and present a similar interface (e.g., PyNN) and between
the underlying neural simulators (e.g., NEST, Brian).

The revised section on neural simulators reads:

"On the neural simulator side, we have shown that both Nengo's
reference simulator and OpenCL simulator are able to simulate two
benchmark models much faster than Brian, NEST and NEURON (see
Figure~\ref{fig:benchmarks}). This is, in part, because Nengo stores
the factors of the connection weight matrix, rather than storing the
entire matrix. However, these simulators are able to simulate many
detailed neuron models and learning rules, and have access to a wealth
of existing neuron models and learning rules. Because Nengo 2.0 is in
an earlier development stage, many of these detailed neuron models and
learning rules remain to be added. Neural simulators like Brian, NEST,
and NEURON are therefore currently better suited for simulating a
wider range of single cell models, while Nengo is designed for large
networks of simple neural models that are connected together according
to the principles of the Neural Engineering Framework, and can
simulate these types of models efficiently.

"One key difference between Nengo's simulators and traditional neural
simulators is the target platform. While NEST and NEURON can be run on
commodity hardware, networks of modest size are typically simulated in
high-performance computing environments by using the Message-Passing
Interface (MPI). Nengo enables large-scale neural simulation on
commodity hardware, allowing researchers and hobbyists without access
to large computing clusters the ability to experiment with theoretical
neuroscience as is currently practiced in cutting edge research. In
particular, GPUs are a powerful, low-cost computing resource that are
available in most modern workstations. The OpenCL simulator makes full
use of GPUs, while the previously discussed simulators currently do
not."

> Where is it available? What do I need in terms of 3rd party software?

Thanks for pointing out that we omitted these important details in the
original manuscript. We have added a link to the release and source
versions of Nengo in a footnote on Page 2 of the revised manuscript
that reads:

"Nengo is available for download through the Python Package Index at
https://pypi.python.org/pypi/nengo. The full source is available at
https://github.com/ctn-waterloo/nengo."

We have also added the following paragraph to the introduction in
order to clarify Nengo's 3rd party dependencies.

"Nengo 2.0 has been rewritten from scratch in Python, leveraging NumPy
(Oliphant, 2007) for manipulating large amounts of data. While NumPy
is its only dependency, Nengo contains optional extensions for
plotting if Matplotlib is available (Hunter, 2007) and for interactive
exploration if IPython 63 available (Perez and Granger, 2007). Since
Nengo only depends on one third-party library, it is easy to integrate
Nengo models in arbitrary CPython programs, opening up possibilities
for using neurally implemented algorithms in web services, games, and
other applications."

> The NEF should be explained in more detail. The paper should be more
> self-contained. It is not entirely clear how the simulations are
> implemented. I think an ensemble of, for example, LIF neurons is
> simulated, and the dynamics and input that they receive AS A
> POPULATION is presented. That is interesting, but I'm not entirely
> sure how this is done. What happens in the Lorenz attractor? Do all
> neurons receive feedback according to the differential equations in
> the script, i.e. as a current? Or as spike trains modulated by this
> system? What is 'input' anyway? This is not described very well at
> all. I would walk the reader through a simple simulation more
> explicitly.

In response to your direct question, all of the neurons in the Lorenz
attractor receive input in the form of current; that current is
determined by a summation of the spiking activities of afferent
neurons weighted by connection weights that are determined using the
NEF's transformation principle.

These details are now included in Appendix 1, "Neural Engineering
Framework Details", which we have added to make this paper more
self-contained. This appendix gives the mathematical details behind
the principles discussed in Section 2. (The text of the Appendix is
too long to include in the interactive review forum.)

We have also described the example scripts in more detail in
sections 5.1, 5.2, and 5.3.

In regard to walking the reader through a simple simulation, we hope
that the addition of the NEF appendix will make the existing Figure 2
a sufficiently detailed account of the first two timesteps of a simple
simulation.

> Having said that, the paper contains far too many irrelevant design
> details. The paragraph on JAVANENGO is bizarre. Do you really want
> to burden you readers with the fact that its class interfaces are
> overloaded? If you want to promote the new Python tool, there is no
> reason to blast the Java version. I recommend dropping the entire
> JAVANENGO section, if the Python version is what you want people to
> use.

Thanks for pointing this out. As the designers of the new version, we
have had many discussions about whether we should improve the old
version or start from scratch; some of that discussion appeared in the
paper, and upon a second look, we wholeheartedly agree that the
mentions of JavaNengo are out of place.

We have removed that section completely in the revised manuscript.
Where mentions of the previous version of Nengo were unavoidable, we
have referred to JavaNengo as Nengo 1.4, and the new version as Nengo
2.0 rather than PyNengo.

> I would mention design considerations only when the relate to use
> cases, i.e. explain why certain scaling properties are achieved. The
> paper contains far too many design considerations, it should focus
> on use cases. Scaling is also an issue of use.
>
> This question is important in order to judge how similar the
> simulator is from other tools such as NEST and BRIAN. Can you set up
> simulations in the two of those that are similar to NENGO and THEN
> compare them? Also bear in mind that NEST has an MPI backbone.

We wanted this paper to focus on the design of Nengo 2.0, and so our
intention in saying that the implemented simulators scale well is
primarily to indicate that the design enables the implementation of
simulators that scale well. It is difficult to come up with use cases
that can be adequately explained in this paper; in a sense, the papers
that are cited in the introduction are those use cases, as they were
all implemented with Nengo 1.4 (Conklin and Eliasmith, 2005; Singh and
Eliasmith, 2006; Choo and Eliasmith, 2010; Rasmussen and Eliasmith,
2014; DeWolf and Eliasmith, 2011; Stewart et al., 2012; Eliasmith et
al., 2012). We believe that the examples discussed in Section 5 are of
sufficient length to be instructive.

Setting up simulations first in Brain or NEST (or PyNN) and then
mapping those to Nengo models would not be very instructive. The point
of Nengo is to build neural models based on the principles of the NEF;
models built in those other tools would not use these principles, and
would result in models containing many ensembles whose underlying
neural populations would be connected directly with a defined weight
matrix. We do make this possible for certain important cases (e.g.,
inhibitory gating signals), but these cases are not the norm when
using the principles of the NEF. We hope that the focus of the paper
is clearer given our additions in the introduction and in Section 7.1.

We agree that it would be an interesting comparison to simulate two
neural populations connected directly with a defined weight matrix.
However, that comparison would be more apt in a paper that gives more
implementation details of the simulators being compared. We have not
provided those details here, and have instead focused on the novel
contributions of Nengo.

In sum, the difference between Nengo and other neural simulators is
primarily that Nengo is designed to define and simulate neural models
created with the principles of the NEF. The updated introduction and
Section 7.1 (copied above) highlights this difference.

> Can you do other neural models than LIF? Again this harks back to
> the criticism that the philosophy of NEF is well described, and
> interesting, but its implementation is not.

Currently in the main development branch of Nengo 2.0, we have only
implemented the LIF model, and a rate approximation of the LIF model.
Nengo 1.4 contained implementations of LIF, an adaptive LIF model
(Koch & Schutter, 1999), and the well-known Hodgkin-Huxley model, so
we know that it is possible to use other neuron models in NEF-based
networks. Additionally, we have already completed work done on another
simulator internal to the lab that uses a neuron model emulating a
piece of neuromorphic hardware.

The added appendix describing the NEF makes explicit how the neuron
model is used. Additionally, we have added the following text to
Section 2:

"Importantly, tuning curves can be determined for any type of neuron,
and therefore the encoding process (and the NEF as a whole) is not
dependent on any particular neuron model."

> Two million spiking neurons is not that much. Edelman and Tononi did
> that 15 years ago. Izhekevich blasts you out of the water. Moreover,
> it is a meaningless quantity, it depends on what you do. 2.5 million
> compartmental neurons is still impressive, 2.5 point model neurons is
> not. In our lab have models with hundreds of thousands of rate-based
> populations, we could model each population by a thousand neurons if
> we wanted to. Does that make us record holder? I'd be a bit more
> careful with such claims, or qualify them better.

We agree that the number of neurons is a meaningless quantity, and
believe that function is the more important metric. That is why our
focus is well-exemplified by Spaun, the largest functional brain model
(which has been previously stated as such in Eliasmith & Trujillo,
2013), where functional in this case refers to the cognitive functions
that Spaun performs.

> In my opinion, this paper cannot be read without going to the original
> paper explaining NEF. Given the brevity of this paper, I think that's
> not good enough. There is ample space for a better explanation of how
> it works and how that is different from more conventional simulators.

Thank you for pointing out these weaknesses. We hope that the addition
of an appendix describing the NEF, an expanded Section 7.1 comparing
Nengo to conventional simulators, and other improvements throughout
the revised manuscript have addressed your concerns and have made this
paper self-contained. In particular, we hope that we have made it
clear that Nengo's strength lies in the fact that it is based on the
principles of the NEF; we do not wish to claim that Nengo is a better
general-purpose neural modeling tool than PyNN, or a better
general-purpose simulator than Brain or NEST. However, we do wish to
claim that Nengo is a better modeling tool and simulator for NEF-style
networks, and that NEF-style networks are less computationally
intensive to simulate, and therefore are feasible to run on commodity
hardware.

Reviewer 2
==========

Expert Authority
----------------

>> Are relevant methods of other authors mentioned? If not, please
>> specify the methods the author should incorporate.

> Authors could perhaps mention the Topographica
> (http://topographica.org/index.html) simulation framework which is
> also implemented in python and offers a high-level model
> specification environment. It also offers a lot of analysis and
> visualization resources and a GUI, which could serve as inspiration
> for some of the future work mentioned in the manuscript.

Thank you for bringing Topographica to our attention! While we have
looked at it some time ago, the project seems to have grown much
lately, and we will investigate it more soon. For now, we have cited
it in Section 7.1, as it seems to have similar goals as Nengo and
PyNN.

Article Length
--------------

>> A Method Article should not exceed 12,000 words. Should any part of
>> the article be shortened? If yes, please specify which part should be
>> shortened.

> I found the introduction unnecessarily lengthy with rather small
> proportion actually discussing the main topic (pyNENGO). Especially
> the JavaNENGO section could in my opinion be omitted or at least
> significantly shortened. Also I found the first part of the
> introduction discussing NEF unnecessarily long - this paper should
> motivate pynengo not NEF.

Thanks for pointing this out. As the designers of the new version, we
have had many discussions about whether we should improve the old
version or start from scratch; some of that discussion appeared in the
paper, and upon a second look, we wholeheartedly agree that the
mentions of JavaNengo are out of place.

We have removed that section completely in the revised manuscript.
Where mentions of the previous version of Nengo were unavoidable, we
have referred to JavaNengo as Nengo 1.4, and the new version as Nengo
2.0 rather than PyNengo.

We have also removed some sections of the introduction motivating the
NEF, and have replaced them with motivations for Nengo. Specifically,
the paragraph:

"It has often been said that neuroscience is data-rich and theory-poor
(Churchland and Sejnowski, 1992). We believe that the modeling success
of large functional networks such as Spaun suggests that the NEF may
be able to fill the theoretical void. In order to continue to evaluate
the NEF as a theory of neural computation by building larger and more
complicated models, we have begun the next generation of Nengo
development with a focus on speed, extensibility, and simplicity."

has been removed, and replace with the following text:

"The transformation principle of the NEF proposes that the connection
weight matrix between two neural populations can compute a nonlinear
function, and can be factored into two significantly smaller matrices.
By using these factors instead of full connection weight matrices,
NEF-designed models are more computationally efficient, which allows
Nengo to run large-scale neural models on low-cost commodity
hardware."

Other Comments
--------------

> Section 1.2 alludes to the possibility of exporting model from pynengo
> to PyNN. Is this a future work, or is there already some
> infrastructure in place for this? Please clarify.

The ability to export a model to PyNN is a future work. We have
removed that confusing passage from Section 1.2 in the revised
manuscript, but that functionality is possible and we hope to
implement it soon.

We have already have two work-in-progress simulators allowing for
Nengo 2.0 models to be run on two pieces of neuromorphic hardware. We
have added a mention of these, copied below, to Section 7.2 in order
to show that multiple backend simulators are possible (including a
future PyNN backend):

"We are also developing two simulators that will take the same NEF
model description as the existing simulators, but will target two
pieces of neuromorphic hardware to achieve greater speed and power
efficiency than the OpenCL simulator."

> In section 3.2 authors discuss parameters, some of which can be set on
> ensembles and some directly on neural populations. It isn't very
> intuitive why some belong to one or the other (i.e. why maximum firing
> rate is a parameter of ensemble rather than the population). I feel it
> could be instructive to discuss a bit more in detail the separation of
> the high level NEF concepts from the low-level spiking neuron concepts
> in pynengo.

The intuition as to why some parameters are ensemble-level and some
are neural population-level relies on a mathematical description of
how we compute the current that we inject into each of the neuron in
the neural population. Any parameter that is part of that equation is
ensemble-level, and any parameter that is not is stored on the neural
population. We have now included these equations in an appendix that
gives a mathematical description of the NEF; specifically, Equation
(1) describes the current that is injected into each neuron model. The
parameters of this equation are stored on the ensemble, whereas the
parameters that are used in the neural nonlinearity would be stored in
the neural population object.

In the revised manuscript we have also removed the mention of the
maximum firing rate in Section 3.1 (formerly 3.2), as the encoding
weight is more intuitively a part of the NEF's representation scheme
specifically and not a property of a neuron.

> I was intrigued by the mention of automatic parameter setting based on
> neurobiological constraints. Could authors discuss this issue further
> in the manuscript?

The description "automatic parameter setting" may be too generous due
to our choice of wording. Specifically, in our original manuscript we
wrote in Section 3.2 (now 3.1):

"If these attributes are not set, Nengo attempts to maintain
neurobiological constraints by selecting neural parameters from
distributions consistent with neocortical pyramidal cells."

However, it is more accurate to say that, if parameters are unset,
Nengo will use defaults that correspond to neocortical pyramidal
cells. The text in the updated manuscript now reads:

"If these attributes are not set, Nengo attempts to maintain
neurobiological constraints by using default parameters consistent
with neocortical pyramidal cells."

We do wish to do some kind of automatic parameters setting in the
future, both in terms of neural parameters and connection parameters.
In the newly added NEF appendix, we also mention how connection
parameters can correspond to experimental data:

"The decoding process is depicted in Figure 1B, where the optimal
linear decoders have been found and used for eight neurons. Notably,
this kind of temporal decoding requires an assumption about the nature
of the temporal filter being used. Here we assume that post-synaptic
currents are such filters, and set the time constants to reflect the
kind of neurotransmitter receptors in the connection (e.g., AMPA
receptors have short time constants, ~10 ms, and NMDA receptors have
longer time constants, ~50 ms)."

> It isn't very clear from the manuscript what Network is. Is it really
> just a simplest form of a container as would appear from the
> description or does it actually offer some distinct functionality?

A Network really is just a simple container, which is used for
organizing large models. Our experience building large models has
shown that this kind of hierarchical organization becomes increasingly
useful as the size of the model increases, and we include Networks for
that reason. The text in Section 3.5 has been revised to explicitly
state that networks are only containers, and can be used for
hierarchical grouping.

> The first paragraph of section 4 implies a simulator dependent
> results. If this is the case, I think it warrants a more detailed
> discussion.

Thanks for pointing this out. We have revised this paragraph to remove
the simulator dependent language, and have moved that information to
Section 4.1, which collects together information specific to the
reference simulator.

The revised first paragraph of Section 4 is:

"Decoupling model creation and simulation has been done previously by
PyNN (Davison et al., 2008). In PyNN, the same Python script can be
used to run a model on four different simulators. Nengo follows this
programming model by decoupling neural model creation and simulation,
which enables Nengo simulators to allocate memory and schedule
computations in the most efficient ways possible. Simulators are given
a Model as an argument; this Model is a static symbolic description.
The simulator can take the model description and build whatever data
structures are best suited to implement the simulation."

> The benchmark results are impressive, but it should be noted that
> most spiking simulators these days are used in HPC environments, and
> thus parallelization comparisons in multi-processor environments
> would be more relevant. If pynengo does not offer this
> functionality, this should be acknowledged and perhaps future plans
> in this direction should be discussed.

This is a good point. While we are currently implementing a simulator
that uses MPI, in this paper we want to highlight that Nengo is able
to simulate large networks that are typically only accessible in HPC
environments on commodity hardware. The paragraph in the introducton,
listed above, makes this explicit early on. We have also included the
following paragraph in Section 7.1:

"One key difference between Nengo's simulators and traditional neural
simulators is the target platform. While NEST and NEURON can be run on
commodity hardware, networks of modest size are typically simulated in
high-performance computing environments by using the Message-Passing
Interface (MPI). Nengo enables large-scale neural simulation on
commodity hardware, allowing researchers and hobbyists without access
to large computing clusters the ability to experiment with theoretical
neuroscience as is currently practiced in cutting edge research. In
particular, GPUs are a rich, low-cost computing resource that are
available in most modern workstations. The OpenCL simulator makes full
use of GPUs, while other simulators currently do not."

> Googling pynengo doesn't return anything, and I thought I can't find
> the package at all until I noticed that in table 1 it says that
> pynengo is the 'nengo' repository. I would advise authors making this
> more clear, to facilitate the promotion of pynengo. Perhaps end the
> introduction or discussion with clear instructions how to access the
> software.
>
> Did authors considered releasing pynengo as pypi package?

Thanks for bringing this up. As mentioned above, we have changed all
references to PyNengo to Nengo 2.0. Additionally, we have pointed to
Nengo on PyPI and Github in a footnote on page 2:

"Nengo is available for download through the Python Package Index at
https://pypi.python.org/pypi/nengo. The full source is available at
https://github.com/ctn-waterloo/nengo."

Currently, we have only registered the nengo namespace on PyPI, but we
are hoping to make an initial 2.0 release before the publication of
this paper.

> Section 7.2 about future talks about future work only in the last
> sentence.

Thank you for noticing this. We have completely rewritten Section 7.2
to more appropriately focus on future work.